---
title: "Explore IMOS Catalogue"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(httr2)
library(dplyr)
library(purrr)
library(lubridate)
```

## Get metadata records from IMOS catalogue

```{r config}
today    <- format(Sys.Date(), "%Y%m%d")
base_url <- "https://catalogue-imos.aodn.org.au/geonetwork/srv/eng/q"

query_params <- list(
  type            = "dataset",
  `_content_type` = "json",
  fast            = "index",
  resultType      = "details"
)
```

```{r fetch-first-page}
fetch_page <- function(from) {
  request(base_url) |>
    req_url_query(!!!query_params, from = from) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)
}

data      <- fetch_page(1)
records   <- data$metadata
n_records <- as.integer(data$summary$`@count`)
cat("Total records:", n_records, "\n")
```

```{r fetch-all-pages}
n_pages <- floor(n_records / 100)

for (i in seq_len(n_pages)) {
  page_data <- fetch_page(i * 100 + 1)
  records   <- c(records, page_data$metadata)
  cat("Page", i + 1, "of", n_pages + 1, "\n")
}

cat("Total records fetched:", length(records), "\n")
```

```{r build-dataframe}
scalar <- function(x) if (is.null(x)) NA_character_ else as.character(x[[1]])

extract_record <- function(rec) {
  info <- rec$`geonet:info`
  list(
    id              = scalar(info$id),
    uuid            = scalar(info$uuid),
    category        = scalar(rec$category),
    type            = scalar(rec$type),
    title           = scalar(rec$title),
    abstract        = scalar(rec$abstract),
    keyword         = list(rec$keyword),
    geoPolygon      = list(rec$geoPolygon),
    geoBox          = list(rec$geoBox),
    tempExtentBegin = scalar(rec$tempExtentBegin),
    tempExtentEnd   = scalar(rec$tempExtentEnd),
    creationDate    = scalar(rec$creationDate)
  )
}

df <- bind_rows(map(records, extract_record))
colnames(df)
```

```{r head}
head(df)
```

## Extract WKT geometries

Spatial extent information in the IMOS catalogue is spread across two fields — `geoPolygon` and `geoBox` — and neither is guaranteed to be present in every record. `geoPolygon` contains the actual polygon geometry as a WKT string or an array of WKT strings when multiple polygons are present. `geoBox` is a simpler bounding box encoded as a `minlon|minlat|maxlon|maxlat` string, used as a fallback when no polygon is available. Some records have both fields, some only one, and some have neither.

To produce a single, consistent `wkt` column we apply the following logic: if `geoPolygon` is present we use it (merging multiple polygons into a `MULTIPOLYGON`); otherwise we fall back to constructing a bounding-box polygon from `geoBox`; if neither field is available the record gets a missing value.

```{r geobox-to-wkt}
geobox_to_wkt <- function(geobox) {
  if (is.null(geobox) || length(geobox) == 0) return(NA_character_)

  # single bbox string: "minlon|minlat|maxlon|maxlat"
  if (is.character(geobox) && length(geobox) == 1) {
    p <- strsplit(geobox, "\\|")[[1]]
    return(sprintf(
      "POLYGON((%s %s, %s %s, %s %s, %s %s, %s %s))",
      p[1], p[2], p[3], p[2], p[3], p[4], p[1], p[4], p[1], p[2]
    ))
  }

  # list of bbox strings (multiple extents)
  if (is.list(geobox)) {
    geobox <- as.character(unlist(geobox))
    combined <- paste(geobox, collapse = "|")

    if (grepl("POINT", combined)) {
      pts <- map_chr(geobox, function(x) {
        p <- strsplit(x, "\\|")[[1]]
        sprintf("(%s %s)", p[1], p[2])
      })
      return(paste0("MULTIPOINT(", paste(pts, collapse = ","), ")"))
    } else {
      polys <- map_chr(geobox, function(x) {
        p <- strsplit(x, "\\|")[[1]]
        sprintf("((%s %s, %s %s, %s %s, %s %s, %s %s))",
                p[1], p[2], p[3], p[2], p[3], p[4], p[1], p[4], p[1], p[2])
      })
      return(paste0("MULTIPOLYGON(", paste(polys, collapse = ","), ")"))
    }
  }

  NA_character_
}
```

```{r extract-wkt}
extract_wkt <- function(geo_polygon, geo_box) {
  # JSON array: list of WKT strings to merge into MULTIPOLYGON
  if (is.list(geo_polygon) && length(geo_polygon) > 0) {
    parts <- map_chr(geo_polygon, function(x) {
      x <- as.character(x)
      x <- gsub("MULTIPOLYGON \\(", "", x)
      x <- gsub("POLYGON", "", x)
      x <- gsub("\\)\\)\\)", "))", x)
      x
    })
    return(paste0("MULTIPOLYGON (", paste(parts, collapse = ","), ")"))
  }

  # missing geoPolygon: fall back to geoBox
  if (is.null(geo_polygon) || length(geo_polygon) == 0) {
    return(geobox_to_wkt(geo_box))
  }

  # single WKT string
  geo_str <- as.character(geo_polygon)
  if (geo_str == "MULTIPOLYGON EMPTY") return(NA_character_)
  geo_str
}

df <- df |>
  mutate(wkt = map2_chr(geoPolygon, geoBox, extract_wkt))
```

## Save keywords

```{r save-keywords}
write.csv(
  data.frame(keyword = map_chr(df$keyword, function(x) {
    if (is.null(x)) NA_character_ else paste(unlist(x), collapse = ";")
  })),
  paste0("keywords_", today, ".csv"),
  row.names = FALSE
)
```

## Calculate temporal extent

```{r temporal-extent}
safe_date <- function(x) tryCatch(as.Date(substr(x, 1, 10)), error = function(e) NA_Date_)

df <- df |>
  mutate(
    temporalExtent = case_when(
      !is.na(tempExtentBegin) & !is.na(tempExtentEnd) ~
        as.numeric(safe_date(tempExtentEnd) - safe_date(tempExtentBegin)) / 365.25,
      .default = NA_real_
    )
  )
```

## Check for wrong temporal extents

```{r wrong-temporal-extent}
df_wrong_extent <- df |> filter(temporalExtent < 0)

if (nrow(df_wrong_extent) > 0) {
  df_wrong_extent |>
    mutate(keyword = map_chr(keyword, function(x) {
      if (is.null(x)) NA_character_ else paste(unlist(x), collapse = ";")
    })) |>
    mutate(across(where(is.list), ~ map_chr(.x, function(x) {
      if (is.null(x)) NA_character_ else paste(unlist(x), collapse = "|")
    }))) |>
    write.csv(paste0("IMOS_wrongTemporalExtent_", today, ".csv"), row.names = FALSE)
}
```

## Save to CSV

```{r save-csv}
df_out <- df |>
  mutate(keyword = map_chr(keyword, function(x) {
    if (is.null(x)) NA_character_ else paste(unlist(x), collapse = ";")
  })) |>
  mutate(across(where(is.list), ~ map_chr(.x, function(x) {
    if (is.null(x)) NA_character_ else paste(unlist(x), collapse = "|")
  })))

write.csv(df_out, paste0("IMOS_datasets_", today, ".csv"), row.names = FALSE)
```
